Reply to reviewers

### 2025-02-04 paper.qmd
Final changes before resubmitting

This approach for comparing new observations with a reference distribution can be applied to any \pD{} problem. @fig-compare illustrates the procedures for a \fD{} problem. The new sample of observations is compared against the reference distribution in two ways (a) against a \fD{} ellipse corresponding to a level curve of density function of the reference distribution and (b) against a sample simulated from the reference distribution. Here we show a single \gD{} projection from a tour of the full \pD{}. The circle and line segments indicate the coefficients of the $4\times 2$ basis used to project the data in these plots. The difference between the new sample of points (dark/red) from the sample is in the bottom left to top right direction, corresponding mostly to the V3 axis direction. Roughly orthogonal to this is a combination of V1 and V2, so the difference between new and reference as indicated by this projection is due to V4 contrasted with V1 and V2. The ellipse makes it clearer see but the sample can be useful, too, particularly if the reference distribution is not a multivariate normal distribution. (The supplementary material includes animated gifs showing these data in a tour, allowing one to see many different projections of the \fD{} points.)

@@

If the observations in $W$ are primarily departing from the normal range in the same direction, the index will be expected to perform well in finding this average direction. However, if the observations have very different departures from the norm, it may be useful to break them into groups, and separately optimize on these subsets. One could consider clustering these observations using angular distance to find groups of observations that have similar directions of departure. This is illustrated in the second example, see Section @sec-weather.

>>>>

This approach for comparing new observations with a reference distribution can be applied to any \pD{} problem. @fig-compare illustrates the procedures for a \fD{} problem. The new sample of observations is compared against the reference distribution in two ways (a) against a \fD{} ellipse corresponding to a level curve of density function of the reference distribution and (b) against a sample simulated from the reference distribution. Here we show a single \gD{} projection from a tour of the full \pD{}. The circle and line segments indicate the coefficients of the $4\times 2$ basis used to project the data in these plots. The difference between the new sample of points (dark/red) from the sample is in the bottom left to top right direction, corresponding mostly to the V3 axis direction. Roughly orthogonal to this is a combination of V1 and V2, so the difference between new and reference as indicated by this projection is due to V4 contrasted with V1 and V2. The ellipse makes it clearer but the sample can be useful, too, particularly if the reference distribution is not a multivariate normal distribution. (The supplementary material includes animated gifs showing these data in a tour, allowing one to see many different projections of the \fD{} points.)

@@

If the observations in $W$ are primarily departing from the normal range in the same direction, the index will be expected to perform well in finding this average direction. However, if the observations have very different departures from the norm, it may be useful to break them into groups, and separately optimize on these subsets. One could consider clustering these observations using angular distance to find groups of observations that have similar directions of departure. This is illustrated in the second example, see @sec-weather.

### 2025-02-04 paper.tex
Final changes before resubmitting

This approach for comparing new observations with a reference
distribution can be applied to any \pD{} problem.
Figure~\ref{fig-compare} illustrates the procedures for a \fD{} problem.
The new sample of observations is compared against the reference
distribution in two ways (a) against a \fD{} ellipse corresponding to a
level curve of density function of the reference distribution and (b)
against a sample simulated from the reference distribution. Here we show
a single \gD{} projection from a tour of the full \pD{}. The circle and
line segments indicate the coefficients of the \(4\times 2\) basis used
to project the data in these plots. The difference between the new
sample of points (dark/red) from the sample is in the bottom left to top
right direction, corresponding mostly to the V3 axis direction. Roughly
orthogonal to this is a combination of V1 and V2, so the difference
between new and reference as indicated by this projection is due to V4
contrasted with V1 and V2. The ellipse makes it clearer see but the
sample can be useful, too, particularly if the reference distribution is
not a multivariate normal distribution. (The supplementary material
includes animated gifs showing these data in a tour, allowing one to see
many different projections of the \fD{} points.)

@@

If the observations in \(W\) are primarily departing from the normal
range in the same direction, the index will be expected to perform well
in finding this average direction. However, if the observations have
very different departures from the norm, it may be useful to break them
into groups, and separately optimize on these subsets. One could
consider clustering these observations using angular distance to find
groups of observations that have similar directions of departure. This
is illustrated in the second example, see Section
Section~\ref{sec-weather}.

@@

\includegraphics[width=2.60417in,height=\textheight,keepaspectratio]{images/anomaly1.png}

@@

\includegraphics[width=2.60417in,height=\textheight,keepaspectratio]{images/anomaly2.png}

@@

\includegraphics[width=2.60417in,height=\textheight,keepaspectratio]{images/liver1.png}

@@

\includegraphics[width=2.60417in,height=\textheight,keepaspectratio]{images/liver2.png}

@@

\includegraphics[width=2.86458in,height=\textheight,keepaspectratio]{images/weather1.png}

@@

\includegraphics[width=2.86458in,height=\textheight,keepaspectratio]{images/weather2.png}

>>>>

This approach for comparing new observations with a reference
distribution can be applied to any \pD{} problem.
Figure~\ref{fig-compare} illustrates the procedures for a \fD{} problem.
The new sample of observations is compared against the reference
distribution in two ways (a) against a \fD{} ellipse corresponding to a
level curve of density function of the reference distribution and (b)
against a sample simulated from the reference distribution. Here we show
a single \gD{} projection from a tour of the full \pD{}. The circle and
line segments indicate the coefficients of the \(4\times 2\) basis used
to project the data in these plots. The difference between the new
sample of points (dark/red) from the sample is in the bottom left to top
right direction, corresponding mostly to the V3 axis direction. Roughly
orthogonal to this is a combination of V1 and V2, so the difference
between new and reference as indicated by this projection is due to V4
contrasted with V1 and V2. The ellipse makes it clearer but the sample
can be useful, too, particularly if the reference distribution is not a
multivariate normal distribution. (The supplementary material includes
animated gifs showing these data in a tour, allowing one to see many
different projections of the \fD{} points.)

@@

If the observations in \(W\) are primarily departing from the normal
range in the same direction, the index will be expected to perform well
in finding this average direction. However, if the observations have
very different departures from the norm, it may be useful to break them
into groups, and separately optimize on these subsets. One could
consider clustering these observations using angular distance to find
groups of observations that have similar directions of departure. This
is illustrated in the second example, see Section~\ref{sec-weather}.

@@

\includegraphics[width=2.60417in,height=\textheight]{images/anomaly1.png}

@@

\includegraphics[width=2.60417in,height=\textheight]{images/anomaly2.png}

@@

\includegraphics[width=2.60417in,height=\textheight]{images/liver1.png}

@@

\includegraphics[width=2.60417in,height=\textheight]{images/liver2.png}

@@

\includegraphics[width=2.86458in,height=\textheight]{images/weather1.png}

@@

\includegraphics[width=2.86458in,height=\textheight]{images/weather2.png}

### 2025-02-04 review/reply.Rmd
Final changes before resubmitting

\textcolor{purple}{We have extended the implementation section to mention the options for optimization of the index, and have included a reference for the ferrn package that can be used to investigate index values and to compare different optimization strategies. All implementations are available in CRAN packages and the code to reproduce the examples are available on GitHub for full reproducibility.}

@@

\textcolor{purple}{We have extended the introduction somewhat to include references to outlier detection in the context of linear dimension reduction. The additional references in the conclusions show potential generalizations of the work connecting to other directionsin outlier detection.}

>>>>

\textcolor{purple}{We have extended the implementation section to mention the options for optimization of the index, and have included a reference for the ferrn package that can be used to investigate index values and to compare different optimization strategies. All implementations are available in CRAN packages and the code to reproduce the examples is available on GitHub for full reproducibility.}

@@

\textcolor{purple}{We have extended the introduction somewhat to include references to outlier detection in the context of linear dimension reduction. The additional references in the conclusions show potential generalizations of the work connecting to other directions in outlier detection.}
